---
title: "Group P Labor Market"
subtitle: "Class Data Visualization"
author: "Ming Nie (mn2984) | Chujun Zhao (cz2621) |Edmund Lam (ecl2169) |Chloe Cao (kc3410)"
date: "2021/04/17"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```


```{r packages, message = FALSE, warning = FALSE, echo=FALSE}

packages <-
  c(
    
    "tidyverse",
    "DT",
    "ggmap",
    "ggthemes",
    "maps",
    "readr",
    "rgeos",
    "tidyverse",
    "plyr",
    "readxl",
    "XML",
    "tmap",
    "leaflet",
    "plotly"
  

  )


packages <- lapply(packages, FUN = function(x) {
  if(!require(x, character.only = TRUE)) {
    install.packages(x)
  library(x, character.only = TRUE)
  }
}
)

```



```{r, message = FALSE, warning = FALSE, echo=FALSE}
df = read.csv("wage.csv")
walmart = read.csv("walmart.csv")
target = read.csv("target.csv")
geo = read.csv("states.csv")
unemploy_2020 <- read_excel("unemploy_2020.xlsx")
```



# Bar Plot: Who is making minimum wage?
```{r , echo=FALSE, warning=FALSE, message=FALSE}
tbl <-list.files(pattern='age_gender*') %>% 
    map_df(~read_csv(.))
tbl <- tbl %>% rename(., `Percentage`=7,
                      `Below minimum wage`=4,
                      `At minimum wage`=5
                      )
tbl$gender <- ifelse(tbl$gender == 'men', 'Men', 'Women')
ag<-tbl %>% ggplot(., aes(x=year, y=`Percentage`, fill=gender, text=paste(Percentage, '% / ', Number,",000", sep='')) ) + 
  geom_col(position='dodge') + 
  facet_wrap(~ age, nrow=1) +
  scale_fill_manual(values=c('blue4','red3'),name = "")+
  labs(x='',y='Percentage', fill='') +
  theme_classic()
p<-ggplotly(ag, tooltip = 'text')%>%
  layout(title= list(text="Percentage of Hourly-paid Workers\n at or below Federal Minimum Wage",font=list(size=15)),
         margin = list(t=100),
         legend=list(orientation='h',x='0.5',xanchor='center',y='-0.1')
         )

p
```

---

# Map 1: State Labor Market General Info Map



```{r, message = FALSE, warning = FALSE, echo=FALSE, out.width = "100%"}
df_2020 = df %>% filter(Year ==2020)
df_2020 <- left_join(df_2020, geo, by=c("State" = "City"))
df_2020 <- left_join(df_2020, unemploy_2020, by=c("State" = "State"))
m <- leaflet(df_2020) %>%
  addTiles() %>%    
  addCircles(lng = ~Longitude, lat = ~Latitude)

content <- paste("State:",df_2020$State,"<br/>",
  "State Minimum Wage",df_2020$State.Minimum.Wage,"<br/>",
                 "Federal Minimum Wage",df_2020$Federal.Minimum.Wage,"<br/>",
                 "Unemployment Rate",df_2020$`Unemployment Rate`,"<br/>")


(mpop2 <- m  %>% addCircleMarkers(color= "#ccffff", weight = 10, popup = content))
```





---

# Map 2: State Minimum Wage Map 2020

```{r, echo=FALSE, out.width = "100%"}
df_2020 <- df %>% filter(Year==2020)
unemploy_2020 <- read_excel("unemploy_2020.xlsx")
# add unemployment data
df_2020 <- left_join(df_2020, unemploy_2020, by=c("State" = "State"))
s <- map_data('state')
df_2020$State <- tolower(df_2020$State)
data <- merge(s, df_2020,
              by.x = 'region',
              by.y = 'State')

ggplot(data, aes(x = long, y = lat, 
                 group = group,
                 fill = State.Minimum.Wage))+labs(caption = "Data Source: National Census Burea") +
  geom_polygon(color = 'gray')+ coord_map('polyconic')+scale_fill_gradient(low = "#ccf8ff", high="blue")
```
---


## Map 3a: What about the Minimum Wage change over the past 20 years?


```{r}
library(tmap)
USA<- map_data("state")
df[[2]] <- tolower(df[[2]])
map_usa <- inner_join(USA, df, by = c("region"="State"))
center <- data.frame(region=tolower(state.name), long=state.center$x, lat=state.center$y)
center$names<-state.abb[match(center$region,tolower(state.name))]
center <- center %>% filter(region!= "alaska") %>% filter(region!= "hawaii") 
```

```{r, echo=FALSE, out.width = "100%"}
library(plotly)
# Minimum Wage Plot for Year 2000
map_usa_2000 <- map_usa%>%filter(Year == 2000)
ggplot()+labs(caption = "Data Source: National Census Burea") + 
  geom_polygon(data=map_usa_2000, aes(x=long, y=lat, group=group, fill = State.Minimum.Wage),
                color="white", size = 0.2) + coord_map('polyconic')+
  geom_text(data=center, aes(long, lat, label = names), size=2)+
  scale_fill_gradient(low = "bisque", high = "firebrick4")  +
  ggtitle("Minimum Wage of Each State in 2000 (in USD)")+
  theme_classic() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) -> map_2000
ggplotly(map_2000, tooltip = "State.Minimum.Wage") 
# On the map: click the label of the state to display the state name, and click the edge of the state area to display minimum wage
```

### Map 3b Minimum Wage in 2008
```{r, echo=FALSE, out.width = "100%"}
# Minimum Wage at the Great Recession
map_usa_2008 <- map_usa%>%filter(Year == 2008)
ggplot() + 
  geom_polygon(data=map_usa_2008, aes(x=long, y=lat, group=group, fill = State.Minimum.Wage),
                color="white", size = 0.2) + coord_map('polyconic')+
  geom_text(data=center, aes(long, lat, label = names), size=2)+
  scale_fill_gradient(low = "bisque", high = "firebrick4") +
  ggtitle("Minimum Wage of Each State at year of the Great Recession (in USD)") +
  theme_classic() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())-> map_2008
ggplotly(map_2008, tooltip = "State.Minimum.Wage")
```

### Map 3c Minimum Wage in 2016
```{r, echo=FALSE, out.width = "100%"}
# Minimum Wage at the Year that Donald Trump was elected
map_usa_2016 <- map_usa%>%filter(Year == 2016)
ggplot() + 
  geom_polygon(data=map_usa_2016, aes(x=long, y=lat, group=group, fill = State.Minimum.Wage),
                color="white", size = 0.2) + coord_map('polyconic')+
  geom_text(data=center, aes(long, lat, label = names), size=2)+
  scale_fill_gradient(low = "bisque", high = "firebrick4") +
  ggtitle("Minimum Wage of Each State in 2016 (in USD)") +
  theme_classic() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())-> map_2016
ggplotly(map_2016, tooltip = "State.Minimum.Wage")
```

### The top 5 States with the Highest Minimum Wage Over Years

```{r,echo=FALSE}
# Extract the top 5 for each year
rank_state <- 
  df %>% 
  select(Year, State, State.Minimum.Wage)%>%
  group_by(Year) %>%  
  arrange(Year, -State.Minimum.Wage) %>%  
  mutate(rank = 1:n()) %>%  
  filter(rank <= 5) 
head(rank_state)
```
```{r, echo=FALSE}
# Initial non-interactive Chart 
wage_prep <- 
  rank_state%>%
  ggplot() +  
  aes(xmin = 0 ,  
      xmax = State.Minimum.Wage) +  
  aes(ymin = rank - 0.45,  
      ymax = rank + 0.45,  
      y = rank) +  
  facet_wrap(~ Year) +
  geom_rect(alpha = 0.5) +
  aes(fill = State) +
    scale_x_continuous(
    limits = c(-16, 40),
    breaks = c(0, 4, 8, 12, 16)) +
  geom_text(hjust = "right",
          aes(label = State),
          x = -1) +
  scale_y_reverse() +
  labs(x = "Minimum Wage",
       y = "") +
  ggtitle("Top 5 States With the Highest State Minimum Wage from 1968 to 2020 (in USD)") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_classic() +
  theme(axis.text.y = element_blank()) +
  theme(axis.ticks.y = element_blank()) +
  theme(axis.line.y = element_blank())+
  theme(legend.title = element_blank())
# Adjust the plot
wage_to_move <- 
wage_prep +  
  facet_null() +
  geom_text(x = 30 , y = -4,
            family = "Garamond",
            aes(label = as.character(Year)),
            size = 15) 
```

```{r fig.height=6}
# Moving chart to show the rank
library(gganimate)
wage_to_move + gganimate::transition_time(Year)
# There are 19 states in the label because only these 19 states made it to the top 5 from 1968 to 2020
```






---

# Map 4a: Sometimes people cannot find jobs 
## Unemployment Rate 2019

```{r, echo=FALSE, out.width = "100%"}
df_2019 <- df %>% filter(Year==2019)
unemploy_2019 <- read_excel("unemploy_2019.xlsx")
unemploy_2019$State <- tolower(unemploy_2019$State)
# add unemployment data
df_2019 <- left_join(df_2019, unemploy_2019, by=c("State" = "State"))
s <- map_data('state')
df_2019$State <- tolower(df_2019$State)
data_2019 <- merge(s, df_2019,
              by.x = 'region',
              by.y = 'State')
data_19 <- rename(data_2019, c( "UnemploymentRate" = "Unemployment Rate" ))

ggplot(data_19, aes(x = long, y = lat, 
                 group = group,
                 fill = UnemploymentRate))+labs(
       caption = "Data Source: National Census Burea") +
  geom_polygon(color = 'gray')+ coord_map('polyconic')+scale_fill_gradient(low = "#fffbcc", high="red")

```

# Map 4b: What about the biggest employer? Where to find a Walmart Shop to work for?

---


```{r, echo = FALSE, fig.align = 'center', message = FALSE}
geo_properties <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showland = TRUE,
  landcolor = toRGB("black"),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)
map2_wal = plotly::plot_geo(walmart, lat = ~latitude, lon = ~longitude, 
                   marker = list(size = 1.7, color = "#fffccc")) %>%
  add_markers(hoverinfo = "none") %>%
  config(displayModeBar = FALSE) %>%
  layout(title = "U.S. Walmart Shop Location Map",geo = geo_properties) 
map2_wal 
         
```



---

# Map 5:  What if I want to work in a Target shop? Target Shop Locations

```{r, message = FALSE, warning = FALSE, echo=FALSE, out.width = "95%"}
#map of shops of target
target <- rename(target, c(  "lat"="Address.Latitude","long"= "Address.Longitude"  ))
content <- paste("State Minimum Wage",target$Name,"<br/>")
factpal <- colorFactor(topo.colors(2), target$SubTypeDescription)
map_target <- leaflet(target) %>%
  addTiles() %>%    
  addCircles(lng = ~long, lat = ~lat, )
(mclust <- map_target %>% addCircleMarkers(color=target$SubTypeDescription, popup = content,clusterOptions = markerClusterOptions()))
```



---


# Scatter Plot 1: Unemployment Rates in All States in 2016
```{r}
# scatter plot in 2016(all states)
cces=read.csv("Unemployall.csv")
#cces <- rename(cces, c( "State" = "ï..State" ))
ggplot(cces, aes(x = State, y = Unemployment.Rate_2016)) +
  geom_point(position = "jitter", alpha=0.5)+
  labs(title="Distribution of unemployment rate in 2016", x="States", y="Unemployment rate") +
  theme(axis.text=element_text(size=6),
        axis.title=element_text(size=6,face="bold"))+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5))+
scale_colour_colorblind()+
guides(size=guide_legend(title=NULL),colour=guide_legend(title=NULL))
```

# Scatter Plot 2: Unemployment Rates in All States in 2020
```{r}
#scatter plot in 2020 all states
ggplot(cces, aes(x = State, y =Unemployment.Rate_2020)) +
  geom_point(position = "jitter", alpha=0.5)+
  labs(title="Distribution of unemployment rate in 2020", x="States", y="Unemployment rate") +
  theme(axis.text=element_text(size=6),
        axis.title=element_text(size=6,face="bold"))+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5))+
  theme(plot.title = element_text(hjust = 0.5))+
scale_colour_solarized()+
guides(size=guide_legend(title=NULL),colour=guide_legend(title=NULL))
```

# Interactive Plot 1: Interactive Plot for the Unemployment Rate Change for Hawaii, New York and Massachusetts

```{r}
#interactive map for Hawaii, New York and Massachusetts
library(plotly)
year<-c(2016,2017,2018,2019,2020)
valueh<-c(2.9,2.2,2.4,2.5,11.6)
valuem<-c(4.0,3.8,3.4,3.0,8.9)
valuen<-c(4.9,4.6,4.1,3.8,10.0)
df <- data.frame(Year = year, Valueh=valueh,Valuem=valuem,Valuen=valuen)
p <- plot_ly(df, x = ~Year, y = ~valueh, name = "Hawaii", type = "scatter", 
    mode = "lines", line = list(color = "rgb(205, 12, 24)", width = 4)) %>% 
    
add_trace(y = ~valuem, name = "Massachusetts", line = list(color = "rgb(22, 96, 167)", 
    width = 4)) %>% 
add_trace(y = ~valuen, name = "New York", line = list(color = "rgb(205, 12, 24)", 
    width = 4, dash = "dash")) %>% 
layout(title = "Unemployment rate in Hawaii, Massachusetts and New York", xaxis = list(title = "Year"), 
    yaxis = list(title = "Unemployment rate"))
p
```


